---
title: "EBC4223 - ASSIGNMENT 1"
author: "Quang Phong - Robert Agatić"
date: "2/14/2022"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = normalizePath("D:/OneDrive/1st-goal-innovation/project-analyze-unstructured-data/3-week3/assignment-1"))
```

# PART 1: INTRODUCTION
The goal of this assignment is to prepare the text data containing ~25 reviews of each of the top 100 movies in 2019 so that we can relate it to movie revenue.

**Import libraries**
```{r, message=FALSE}
library(tm)
library(qdap)
library(SnowballC)
library(dplyr)
library(PerformanceAnalytics)
```

**Import data**
```{r}
load("Data movie reviews final.RData")
```

**Have a look of the data for movies**
```{r}
# rename the data frame
df_Movies <- moviedescriptives 

# get an overview of the first 6 movies
head(df_Movies)
```
```{r}
# get an overview of variables
str(df_Movies)
```
This data set is about `r colnames(df_Movies)` of 100 movies.

**Have a look of the data set for reviews**
```{r}
# rename the data frame
df_Reviews <- reviews 

# get an overview of the first 6 movies
head(df_Reviews, 1)
```
```{r}
summary(df_Reviews)
```
```{r}
dim(df_Reviews)
```
This data set is about `r colnames(df_Reviews)` of 2469 reviews.


# PART 2: DATA PREPROCESSING


**Ensure the review text is in text format**
```{r}
df_Reviews$review_text <- as.character(df_Reviews$review_text)
```

**Get an overview of the first 5 reviews**
```{r}
df_Reviews$review_text[1:5]
```


**Obtain the length of each review and add it to review data set**
```{r}
df_Reviews$length <- nchar(df_Reviews$review_text)
head(df_Reviews$length)
```
**Describe and plot the amount of characters**
```{r}
summary(df_Reviews$length)
hist(df_Reviews$length)
boxplot(length~movie_id, data = df_Reviews, main = "Characters per movie reviews",
        xlab = "Movie ID", ylab = "Number of characters")
```


**Obtain average review length for each movie and add it to movie data set**
```{r}
tempdf_ReviewLength <- df_Reviews %>%
  group_by(Release) %>%
  summarize(averageReviewLength = mean(length))
head(tempdf_ReviewLength)
```
```{r}
# Merge the average review length to df_Movies
df_Movies <- df_Movies %>%
  inner_join(tempdf_ReviewLength, by = "Release")

head(df_Movies)
```

**Remove punctuations from review data**
```{r}
df_Reviews$review_text2 <- removePunctuation(df_Reviews$review_text)
head(df_Reviews$review_text2, 1)
```
**Remove numbers from review data**
```{r}
df_Reviews$review_text2 <- removeNumbers(df_Reviews$review_text2)
head(df_Reviews$review_text2, 1)
```
**Lower texts from review data**
```{r}
df_Reviews$review_text2 <- tolower(df_Reviews$review_text2)
head(df_Reviews$review_text2, 1)
```

**Remove excess white space from review data**
```{r}
df_Reviews$review_text2 <- stripWhitespace(df_Reviews$review_text2)
head(df_Reviews$review_text2, 1)
```

**Remove potentially problematic symbols from review data**
```{r, results='hide'}
# apply(df_Reviews['review_text'], 1, function(x) gsub("[[:punct:]]", "", x))
df_Reviews$review_text2 <- gsub("[[:punct:]]", "", df_Reviews$review_text2)
```

**Create variable indicating the amount of words**
```{r}
df_Reviews$amount_words <- sapply(strsplit(df_Reviews$review_text2, " "), length)
```

**Remove stop words from review data**
```{r}
# Here, we choose to remove stop words in the built-in dictionary. 
df_Reviews$review_text2 <- removeWords(df_Reviews$review_text2, stopwords("en"))
head(df_Reviews$review_text2, 1)

# Another choice to remove all stop words in the dictionary except for "not", by:
# stop_words <- stop_words("english")
# stop_words <- stop_words(-which(stop_words=="not"))
```

**Stem review data**
```{r}
df_Reviews$review_text2 <- stemDocument(df_Reviews$review_text2, language = "english")
head(df_Reviews$review_text2, 1)
```


# PART 3: ANALYSIS


**Calculate polarity scores for each individual review**
```{r}
df_Reviews$polarity <- as.numeric(counts(polarity(df_Reviews$review_text2))[, "polarity"]) 
# apart from qdap, we can also use sentimentR library
```

**Describe and plot the polarity scores**
```{r}
summary(df_Reviews$polarity)
hist(df_Reviews$polarity)
boxplot(polarity~movie_id, data = df_Reviews,
        main = "Polarity score per movie review",
        xlab = "Movie ID",
        ylab = "Polarity Score")
```


**Obtain average review polarity, standard deviation and add them to movie data**
```{r}
tempdf_ReviewPolarity <- df_Reviews %>%
  group_by(Release) %>%
  summarize(reviewPolarity = mean(polarity),
            stdPolarity = sd(polarity))
head(tempdf_ReviewPolarity)
```

```{r}
# Merge the average polarity to df_Movies
df_Movies <- df_Movies %>%
  inner_join(tempdf_ReviewPolarity, by = "Release")

head(df_Movies)
```

**Run regression models to explain revenue variable**

```{r}
model1 <- lm(Gross ~ reviewPolarity, data = df_Movies)
model2  <- lm(Gross ~ reviewPolarity+stdPolarity, data = df_Movies)
model3 <- lm(Gross ~ reviewPolarity+stdPolarity+averageReviewLength, data = df_Movies)
model4 <- lm(Gross ~ reviewPolarity+stdPolarity+averageReviewLength+`Max Th`, data = df_Movies)
```

```{r}
summary(model1)
summary(model2)
summary(model3)
summary(model4)
```
**Run regression models with ln of revenue variable as outcome variable**
```{r}
df_Movies$lnGross <- log(df_Movies$Gross)
```


```{r}
model5 <- lm(lnGross ~ reviewPolarity, data = df_Movies)
model6  <- lm(lnGross ~ reviewPolarity+stdPolarity, data = df_Movies)
model7 <- lm(lnGross ~ reviewPolarity+stdPolarity+averageReviewLength, data = df_Movies)
model8 <- lm(lnGross ~ reviewPolarity+stdPolarity+averageReviewLength+`Max Th`, data = df_Movies)
```

```{r}
summary(model5)
summary(model6)
summary(model7)
summary(model8)
```
**Save and restore my entire workspace**
```{r}
save.image(file = "assignment1.RData")
saveRDS(df_Reviews,"df_Reviews.RDS")
saveRDS(df_Movies,"df_Movies.RDS")
```

